Quantitative Benchmarking for Automatically Generated Conjectures

### 2019-02-08 TEBenchmarkPaper/article.tex
Embed CSV data in .tex file
Embed all files except figures in latex
Try to address sampling points

% Derived from the template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
\usepackage{csvsimple}
\usepackage[inline, shortlabels]{enumitem}
\usepackage{epstopdf}
\usepackage{hyperref}
\usepackage{mathtools}
\usepackage{tikz}
\usetikzlibrary{arrows, calc, decorations.pathreplacing, positioning,
  shapes.geometric}
% etc.

@@

To solve both of these problems we instead \emph{sample} a subset of
definitions.\footnote{This could be done randomly, but for reproducibility we
  use a deterministic order based on cryptographic hashes.} Given a sample size,
we choose a subset of that many definitions, and provide only those as input to
the tool. We generate a corresponding ground truth by selecting those properties
from the corpus which ``depend on'' (contain references to) \emph{only} the
definitions in that sample. Transitive dependencies aren't required (e.g. a
property involving only a \texttt{times} function would not depend on a
\texttt{plus} function, even if \texttt{plus} occurs in the definition of
\texttt{times}).

Unfortunately, uniform sampling of definitions gives rise to a lottery: for a
given sample size, increasing the size of the corpus (which provides better
statistics) makes it less likely that a chosen sample will contain all of
a property's dependencies. The majority of such samples would hence have an
empty ground truth, and thus $0$ precision and undefined recall
\emph{independent} of the tool's output! This is clearly undesirable as an
evaluation method.

Other measurements could be used for such cases, but our approach is to avoid
them by only allowing a sample if it contains all dependencies of at least one
property. We could do this using rejection sampling, but it is more efficient to
pick a property, weighted in proportion to their number of dependencies
(ignoring those with more dependencies than our sample size). That property's
dependencies become our sample, padded up to the required size with uniform
choices from the remaining definitions.

The ground truth for such samples is guaranteed to contain at least one property
(the one we picked), and hence the precision and recall will depend meaningfully
on the tool's output. One downside of this restriction is that we limit the
number of possible samples to measure; this is significant for sample sizes less
than~3.

>>>>

% Derived from the template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}

%INCLUDE STYLES HERE

%INCLUDE SPEEDUP CSV HERE

%INCLUDE BIBTEX HERE

\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
\usepackage{csvsimple}
\usepackage[inline, shortlabels]{enumitem}
\usepackage{epstopdf}
\usepackage{hyperref}
\usepackage{mathtools}
\usepackage{tikz}
\usetikzlibrary{arrows, calc, decorations.pathreplacing, positioning,
  shapes.geometric}
% etc.

@@

To solve both of these problems we instead \emph{sample} a subset of
definitions.\footnote{This could be done randomly, but for reproducibility we
  use a deterministic order based on cryptographic hashes.} The sample size is a
free parameter so that users have some control over how long the runs will take
($\S$\ref{sec:application} shows that larger samples take longer, but the
correlation is very noisy). Given a sample size, we need to choose a subset of
that many definitions from the corpus, and provide only those as input to the
tool.

We cannot expect an MTE tool to produce all of the properties found in the
corpus, if it is only given a subset of definitions as input. Hence for each
sample we generate a corresponding ground truth by selecting those properties
from the corpus which ``depend on'' (contain references to) \emph{only} the
definitions in that sample. Transitive dependencies aren't required (e.g. a
property involving only a \texttt{times} function would not depend on a
\texttt{plus} function, even if \texttt{plus} occurs in the definition of
\texttt{times}).

The question remains of how to pick definitions for each sample. One simple way
is to choose uniformly without replacement, but this will give many samples
whose corresponding ground truths are empty. The effect is similar to a lottery,
where our sample is a ticket and must contain all of some property's
dependencies in order to ``win'' (i.e. contain that property in its ground
truth). In particular, if we choose a sample of size $S$ uniformly at random
from a corpus containing $N$ definitions and $P$ properties which depend on an
average of $D$ definitions each (assuming no properties have the exact same
dependencies), then the probability that our sample contains the dependencies of
some particular property $p$ follows a hypergeometric distribution (similar to
the binomial distribution, but without replacement):

\begin{equation*}
  Pr(p) = \frac{\binom{S}{D} \binom{N - S}{S - D}}
               {\binom{N}{S}}
\end{equation*}

Choosing at least one property's dependencies is a Bernoulli process with $P$
repetitions, with probability $1 - (1 - Pr(p))^P$. This gets larger if we
increase the sample size $S$ or the number of properties $P$, but it gets
smaller if we increase the number of definitions $N$ or the average number of
dependencies $D$. This is problematic because we would like large corpora for
better statistics, but this would make empty ground truths more likely.

Samples whose corresponding ground truth is empty are problematic because they
have $0$ precision and undefined recall (due to division by $0$). Such samples
are hence unsuitable for our benchmark, especially since these results would be
\emph{independent} of the tool's output (all tools would fare the same, so no
comparisons could be made). Other measurements could be used for such cases, but
since our choice of analysis comes from existing work we would rather avoid such
samples by only allowing those with a non-empty ground truth (i.e. containing
all dependencies of at least one corpus property).

We can achieve this by instead performing uniform sampling \emph{with
  rejection}, where samples with empty ground truth are rejected and a different
sample is drawn instead.

One problem with performing rejection sampling is that it becomes very
inefficient when the probability of rejection is high. In our case there is a
more direct method of sampling, whose output is exactly equivalent to rejection
sampling but is far more efficient.

Since non-rejected samples must contain the dependencies of some property, we
can consider each such sample to be the union of those dependencies and some
other ``padding'' definitions. By choosing these two components separately and
returning their union, we can obtain an identical distribution without having to
perform any rejection.

We first choose the set of dependencies by picking from those of all corpus
properties, weighted according to the probability that they would appear under
uniform sampling. To do this we first discard any dependency sets which are
larger than the given sample size (since they cannot appear as subsets); we then
discard any dependency set which is a superset of another (since only the subset
is required to avoid rejection). Their probabilities under uniform sampling
depend only on their size, so we assign weights by calculating the least
common multiple (LCM) of their sizes, and dividing that by each set's size. For
example, given dependency sets $X = \{a, b, c\}$, $Y = \{b, d\}$ and
$Z = \{a, b, e, f\}$ the LCM of their lengths ($3$, $2$ and $4$) is 12, so the
weighting will be:

\begin{equation*}
  w(X) = \frac{12}{3} = 4
  w(Y) = \frac{12}{2} = 6
  w(Z) = \frac{12}{4} = 3
\end{equation*}

Once we've chosen a dependency set, we know how much padding to choose (sample
size minus dependency set size). The padding will be uniformly distributed,
since it does not affect the rejection criterion.

Using rejection sampling or, equivalently, our direct method, ensures that the
analysis is well defined and depends meaningfully on the tool's output. One
downside of applying this non-empty ground truth restriction is that we limit
the number of possible samples to measure; this is significant for sample sizes
less than~3.
